---
title: "Dplyr warm-up - úkoly ze začátku semináře 27-10-2020"
output:
  html_document:
    code_foldin: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

countries = read.csv(here("data", "countries.csv"))
```


### 1. úkol

* 1.1 select variables country, uni_prc, life_exp and only observations 
* 1.2 do the same but only for countries with dem_index bigger than 6
* 1.3 and last time the same, but only for countries with dem_index below median dem_index and show from highest life_exp
* 1.4 how can we find the one which has median dem_index?

Pro odpověď klikněte na tlačítko **code**:
```{r question 1, eval=FALSE}
countries %>% select(country, uni_prc, life_exp) # 1.1

countries %>% filter(dem_index > 6) %>% select(country, uni_prc, life_exp) # 1.2

countries %>% filter(dem_index < median(dem_index, na.rm = TRUE)) %>% 
  select(country, uni_prc, life_exp) %>% arrange(desc(life_exp)) #1.3

countries %>% filter(dem_index == median(dem_index, na.rm = TRUE)) #1.4
```

### 2. úkol

* 2.1 what is the max dem_index and min dem_index?
* 2.2 what is the max and min dem_index in each maj_belief category

Pro odpověď klikněte na tlačítko **code**:
```{r question 2, eval=FALSE}

countries %>% summarise(maximum = max(dem_index, na.rm = TRUE),
                        minimum = min(dem_index, na.rm = TRUE)) #2.1

countries %>% group_by(maj_belief) %>%
  summarise(maximum = max(dem_index, na.rm = TRUE), minimum = min(dem_index, na.rm = TRUE)) #2.2
```

### 3. úkol

* 3.1 create new variable - uni_prc_100
* 3.2 create new variable - difference from mean life_exp within post_soviet

Pro odpověď klikněte na tlačítko **code**:
```{r question 3, eval=FALSE}
countries %>% mutate(uni_prc_100 = uni_prc*100) %>%
  select(code, uni_prc, uni_prc_100) # 3.1

countries %>% group_by(postsoviet) %>% 
  mutate(life_exp_within_postsovier = mean(life_exp, na.rm=TRUE),
         life_exp_diff_within_postsoviet = life_exp - mean(life_exp, na.rm = TRUE)) %>% 
  select(code, life_exp,life_exp_within_postsovier, life_exp_diff_within_postsoviet) # 3.2
```
